---
title: "Datacleaning again"
author: "Isabelle Huynh 540706002"
date: "2026-09-03"
output: html_document
---

```{r}
library(dplyr)
library(readr)
library(dplyr)
library(readr)
library(stringr)

files <- list.files(
  "~/Desktop/STAT3888 Project 1/NHS 11-12", # Modify the location
  pattern = "^AHS.*\\.csv$", # # Regex : modify this to filter different files
  full.names = TRUE
)
variable_catalogue <- data.frame()

for (file in unique(files)) {
  
  vars <- names(
    read_csv(
      file,
      n_max = 0, # Read column names
      show_col_types = FALSE
    )
  )
  
  temp <- data.frame( # temporary data frame for one file
    File = basename(file),
    Variable = vars
  )
  
  variable_catalogue <- bind_rows(
    variable_catalogue,
    temp
  )
}
variable_catalogue %>%
  filter(Variable == "ALCUSUQ2") # alc frequency in last 12 months 
  filter(Variable == "FEMLSBC")

```

```{r}

# Read in the two CSV files
df1 <- read_csv("~/Desktop/STAT3888 Project 1/NHS 11-12/AHSnhs11bsp.csv")   
                # e.g. contains ABSLID, sex
df2 <- read_csv("~/Desktop/STAT3888 Project 1/NHS 11-12/AHSnhs11bbi.csv")   
# e.g. contains ABSLID, chronic kidney disease status


# Remove the "NHS11B" prefix from ABSLID in both files
df1 <- df1 %>%
  mutate(ABSLID = str_remove(ABSLID, "^NHS11B"))

df2 <- df2 %>%
  mutate(ABSLID = str_remove(ABSLID, "^NHS11B"))

# Now merge as before
merged_df <- inner_join(df1, df2, by = c("ABSLID", "ABSPID"))

# Select just the columns you need (adjust column names to match your data)
selected_df <- merged_df %>%
  select(ABSLID, SEX, EGFRKDSA, ACRNTR, ALCUSUQ2, FEMLSBC)   # replace 'sex' and 'ckd_status' with your actual column names

# Count how many people of each sex have CKD vs don't
sex_ckd_counts <- selected_df %>%
  group_by(EGFRKDSA, SEX ) %>%
  summarise(n = n(), .groups = "drop")

menopause_counts <- selected_df %>%
  group_by(FEMLSBC) %>%
  summarise(n = n(), .groups = "drop")

print(sex_ckd_counts)

#Data cleaning
#There is a duplicate of Row 12909/12910 with ID NHS11B0000769
# NHS11B0004830 appears to 
```

```
```{r}
library(ggplot2)
theme_set(theme_classic())

# Histogram on a Continuous (Numeric) Variable
g <- ggplot(selected_df, aes(EGFRKDSA)) + scale_fill_brewer(palette = "Spectral")

g + geom_histogram(aes(fill=SEX), 
                   binwidth = .1, 
                   col="black", 
                   size=.1) +  # change binwidth
  labs(title="Histogram with Auto Binning", 
       subtitle="Engine Displacement across Vehicle Classes")  

g + geom_histogram(aes(fill=SEX), 
                   bins=5, 
                   col="black", 
                   size=.1) +   # change number of bins
  labs(title="Histogram with Fixed Bins", 
       subtitle="Engine Displacement across Vehicle Classes") 
```


