---
title: "logistic_Yilan"
output: html_document
date: "2026-09-24"
---

```{r}
library(dplyr)
library(emmeans)
library(car)
library(dagitty)
```

```{r}
ckd_filter <- readRDS("shared_df_filter.rds")
```

```{r}
class(ckd_filter$Cider)
```

```{r}
table(ckd_filter$AGE, useNA = "ifany") 
```
"Other" was excluded because all observations had a value of 0 and therefore provided no variation for the analysis.

The “No alcohol (past year)” and “Never alcohol” categories contained no participants (n = 0) and were excluded from the regression analysis.

```{r}
# number of people by number of alcohol types consumed
alcohol_types <- c(
  "Low_alcohol_beer",
  "Mid_strength_beer",
  "Full_strength_beer",
  "White_wine",
  "Red_wine",
  "Low_alcohol_wine",
  "Champagne_sparkling_wine",
  "Spirits_RTD",
  "Liqueurs",
  "Spirits",
  "Fortified_wine",
  "Cider",
  "Cocktails"
)

ckd_filter <- ckd_filter %>%
  mutate(
    n_alcohol_types = rowSums(
      across(all_of(alcohol_types))
    )
  )

table(ckd_filter$n_alcohol_types)
```

## Causal Analysis
```{r}
dag_alcohol_ckd <- dagitty("
dag {
AGE -> Alcohol
AGE -> CKD

SEX -> Alcohol
SEX -> CKD

Alcohol -> CKD
}
")

adj_alcohol_ckd <- adjustmentSets(
  dag_alcohol_ckd,
  exposure = "Alcohol",
  outcome = "CKD"
)

adj_alcohol_ckd
```


### Full model
```{r}
# Using the original predictors, full model with interaction
full_model <- glm(CKD ~ Alcohol_risk * SEX_factor + alcohol_quantity * SEX_factor + AGE + ALC_freq_3_factor + Low_alcohol_beer + Mid_strength_beer + Full_strength_beer + White_wine + Red_wine + Low_alcohol_wine + Champagne_sparkling_wine + Champagne_sparkling_wine + Spirits_RTD + Liqueurs + Fortified_wine + Cider + Cocktails, data = ckd_filter, family = binomial)
summary(full_model)
```

```{r}
vif(full_model)
```


```{r}
# Do not run this
# Combine(When alcohol types are numeric!)
ckd_filter$Beer <- as.integer(
  rowSums(ckd_filter[, c(
    "Low_alcohol_beer",
    "Mid_strength_beer",
    "Full_strength_beer"
  )], na.rm = TRUE) > 0
)

ckd_filter$Wine <- as.integer(
  rowSums(ckd_filter[, c(
    "White_wine",
    "Red_wine",
    "Low_alcohol_wine",
    "Champagne_sparkling_wine",
    "Fortified_wine"
  )], na.rm = TRUE) > 0
)

ckd_filter$Spirits_Liqueurs <- as.integer(
  rowSums(ckd_filter[, c(
    "Spirits_RTD",
    "Liqueurs",
    "Spirits"
  )], na.rm = TRUE) > 0
)

ckd_filter$Other <- as.integer(
  rowSums(ckd_filter[, c(
    "Cider",
    "Cocktails"
  )], na.rm = TRUE) > 0
)

```

```{r}
ckd_filter$Beer <- factor(
  as.character(ckd_filter$Low_alcohol_beer) == "1" |
  as.character(ckd_filter$Mid_strength_beer) == "1" |
  as.character(ckd_filter$Full_strength_beer) == "1",
  levels = c(FALSE, TRUE),
  labels = c("0", "1")
)

ckd_filter$Wine <- factor(
  as.character(ckd_filter$White_wine) == "1" |
  as.character(ckd_filter$Red_wine) == "1" |
  as.character(ckd_filter$Low_alcohol_wine) == "1" |
  as.character(ckd_filter$Champagne_sparkling_wine) == "1" |
  as.character(ckd_filter$Fortified_wine) == "1",
  levels = c(FALSE, TRUE),
  labels = c("0", "1")
)

ckd_filter$Spirits_Liqueurs <- factor(
  as.character(ckd_filter$Spirits_RTD) == "1" |
  as.character(ckd_filter$Liqueurs) == "1" |
  as.character(ckd_filter$Spirits) == "1",
  levels = c(FALSE, TRUE),
  labels = c("0", "1")
)

ckd_filter$Other <- factor(
  as.character(ckd_filter$Cider) == "1" |
  as.character(ckd_filter$Cocktails) == "1",
  levels = c(FALSE, TRUE),
  labels = c("0", "1")
)
```


```{r}
table(ckd_filter$AGE, useNA = "ifany")
table(ckd_filter$SEX_factor, useNA = "ifany")
table(ckd_filter$Beer, useNA = "ifany")
table(ckd_filter$Wine, useNA = "ifany")
table(ckd_filter$Spirits_Liqueurs, useNA = "ifany")
table(ckd_filter$Other, useNA = "ifany")
```


Note: Other is not a variable from the original dataset. It represents cider or cocktail consumption.

### Only quantity
```{r}
# Using combined predictors, only quantity
model_quantity_int <- glm(CKD ~ AGE + alcohol_quantity * SEX_factor + Beer + Wine + Spirits_Liqueurs + Cider + Cocktails, data = ckd_filter, family = binomial)
summary(model_quantity_int)
```

Association: using logistic regression doesn't need to handle class imbalance?
interaction p: >0.05 but very close?


```{r}
model_quantity_no_int <- glm(CKD ~ alcohol_quantity + SEX_factor + Beer + Wine + Spirits_Liqueurs + Cider + Cocktails, data = ckd_filter, family = binomial)
summary(model_quantity_no_int)
```

```{r}
anova(model_quantity_int,model_quantity_no_int, test="Chisq")
```

```{r}
AIC(model_int_no_risk,model_no_risk)
```

The likelihood-ratio test indicated that adding the interaction term significantly improved model fit, while the interaction model also had a slightly lower AIC

```{r}
anova(model_int_no_risk)
```

### Only risk levels
```{r}
# Risk no quantity
model_risk_no_int <- glm(
  CKD ~ Alcohol_risk + SEX_factor +
        Beer + Wine + Spirits_Liqueurs + Cider + Cocktails,
  data = ckd_filter,
  family = binomial
)

model_risk_int <- glm(
  CKD ~ Alcohol_risk * SEX_factor +
        Beer + Wine + Spirits_Liqueurs + Cider + Cocktails,
  data = ckd_filter,
  family = binomial
)
```

```{r}
summary(model_risk_int)
```

```{r}
summary(model_risk_no_int)
```

```{r}
anova(
  model_risk_no_int,
  model_risk_int,
  test = "Chisq"
)
```




```{r}
vif(model_no_int)
```



### Quantity + risk levels
```{r}
# Quantity + risk
model_int <- glm(CKD ~ alcohol_quantity * SEX_factor + Alcohol_risk * SEX_factor + Beer + Wine + Spirits_Liqueurs + Other, data = ckd_filter, family = binomial)
summary(model_int)
```

```{r}
model_no_int <- glm(CKD ~ alcohol_quantity + Alcohol_risk + SEX_factor + Beer + Wine + Spirits_Liqueurs + Other, data = ckd_filter, family = binomial)
summary(model_no_int)
```

```{r}
vif(model_no_int)
```
No major multicollinearity!

### Compare male and female
```{r}
# Male and female
model_male <- glm(
  CKD ~ alcohol_quantity + Alcohol_risk + Beer + Wine + Spirits_Liqueurs + Other,
  data = subset(ckd_filter, SEX_factor == "Male"),
  family = binomial
)

model_female <- glm(
  CKD ~ alcohol_quantity + Alcohol_risk + Beer + Wine + Spirits_Liqueurs + Other,
  data = subset(ckd_filter, SEX_factor == "Female"),
  family = binomial
)

summary(model_female)
summary(model_male)
```

```{r}
emm <- emmeans(
  model_int,
  ~ Alcohol_risk | SEX_factor
)

pairs(
  emm,
  type = "response",
  infer = c(TRUE, TRUE)
)
```

Compare the odds of CKD between different alcohol-risk groups within each sex.  
Among females, Low risk differed significantly from High risk, no significant differences were observed for the other comparisons.  
Among males, Low risk was associated with higher CKD odds than Risky and High risk.  
p: Assuming 2 groups have no differences, the probability of observing this result

OR>1, odds of the previous group is higher -> higher probability of no CKD

## Model Selection
interaction or no interaction
include risk?
...We don't need to do model selection because we can only choose the model that's related to our RQ (with interaction, consider quantity or risk)?

## Model Performance


