---
title: "logistic_Yilan"
output: html_document
date: "2026-09-24"
---

```{r}
library(dplyr)
```

```{r}
ckd_filter <- readRDS("shared_df_filter.rds")
```


```{r}
class(ckd_filter$Alcohol_risk_7day)
```
```{r}
table(ckd_filter$Other, useNA = "ifany") # 
```
"Other" was excluded because all observations had a value of 0 and therefore provided no variation for the analysis.

```{r}
# number of people by number of alcohol types consumed
alcohol_types <- c(
  "Low_alcohol_beer",
  "Mid_strength_beer",
  "Full_strength_beer",
  "White_wine",
  "Red_wine",
  "Low_alcohol_wine",
  "Champagne_sparkling_wine",
  "Spirits_RTD",
  "Liqueurs",
  "Spirits",
  "Fortified_wine",
  "Cider",
  "Cocktails"
)

ckd_filter <- ckd_filter %>%
  mutate(
    n_alcohol_types = rowSums(
      across(all_of(alcohol_types))
    )
  )

table(ckd_filter$n_alcohol_types)
```

```{r}
# Using the original predictors, full model with interaction
res <- glm(CKD ~ alcohol_quantity * SEX_factor + Low_alcohol_beer + Mid_strength_beer + Full_strength_beer + White_wine + Red_wine + Low_alcohol_wine + Champagne_sparkling_wine + Champagne_sparkling_wine + Spirits_RTD + Liqueurs + Fortified_wine + Cider + Cocktails, data = ckd_filter, family = binomial)
summary(res)
```


```{r}
# Combine
ckd_filter$Beer <- as.integer(
  rowSums(ckd_filter[, c(
    "Low_alcohol_beer",
    "Mid_strength_beer",
    "Full_strength_beer"
  )], na.rm = TRUE) > 0
)

ckd_filter$Wine <- as.integer(
  rowSums(ckd_filter[, c(
    "White_wine",
    "Red_wine",
    "Low_alcohol_wine",
    "Champagne_sparkling_wine",
    "Fortified_wine"
  )], na.rm = TRUE) > 0
)

ckd_filter$Spirits_Liqueurs <- as.integer(
  rowSums(ckd_filter[, c(
    "Spirits_RTD",
    "Liqueurs",
    "Spirits"
  )], na.rm = TRUE) > 0
)

ckd_filter$Other <- as.integer(
  rowSums(ckd_filter[, c(
    "Cider",
    "Cocktails"
  )], na.rm = TRUE) > 0
)

```

Note: Other is not a variable from the original dataset. It represents cider or cocktail consumption.

```{r}
# Using combined predictors
model_interaction <- glm(CKD ~ alcohol_quantity * SEX_factor + Beer + Wine + Spirits_Liqueurs + Other, data = ckd_filter, family = binomial)
summary(model_interaction)
```

Association: using logistic regression doesn't need to handle class imbalance?

interaction p: >0.05 but very close?

```{r}
model_no_interaction <- glm(CKD ~ alcohol_quantity + SEX_factor + Beer + Wine + Spirits_Liqueurs + Other, data = ckd_filter, family = binomial)
summary(model_no_interaction)
```

```{r}
anova(model_no_interaction,model_interaction, test="Chisq")
```

```{r}
AIC(model_no_interaction,model_interaction)
```

The likelihood-ratio test indicated that adding the interaction term significantly improved model fit, while the interaction model also had a slightly lower AIC

```{r}
anova(model_interaction)
```


## Model Selection
stepwise?

## Model Performance
