---
title: "random_forest"
output: html_document
---

```{r}

library(caret)
library(rpart)
library(rpart.plot)
library(tidyverse)
library(randomForest)
library(dplyr)
library(ranger)
library(pROC)
library(ggplot2)
library(pdp)

set.seed(123)

df_filter <- readRDS("shared_df_filter.rds")


bev_cols <- c("Low_alcohol_beer","Mid_strength_beer","Full_strength_beer",
              "White_wine","Red_wine","Low_alcohol_wine",
              "Champagne_sparkling_wine","Spirits_RTD","Liqueurs",
              "Spirits","Fortified_wine","Cider","Cocktails")

rf_data <- df_filter %>%
  select(-SEX, -Alcohol_risk, -ALC_freq_3) 

table(train$CKD) / nrow(train)   # check class imbalance

#Need to fix class imbalance

idx   <- sample(nrow(rf_data), 0.7 * nrow(rf_data))
train <- rf_data[idx, ]
test  <- rf_data[-idx, ]

train_maj <- train %>% filter(CKD == "0")
train_min <- train %>% filter(CKD == "1")


train_up <- bind_rows(
  train_maj,
  train_min[sample(nrow(train_min), nrow(train_maj), replace = TRUE), ]
)

table(train_up$CKD)   # should now be balanced

rf <- randomForest(CKD ~ .,
                  data = train_up,
                  ntree = 500,
                  importance = TRUE,
 )

rf

# Evaluate on the untouched test set
pred_class <- predict(rf, test)
pred_prob  <- predict(rf, test, type = "prob")[, "1"]

table(Predicted = pred_class, Actual = test$CKD)
auc(roc(test$CKD, pred_prob))


tree <- rpart(CKD ~ ., data = train_up, method = "class",
              control = rpart.control(cp = 0.005, minbucket = 30))

```

```{r}
mda_scores <- importance(rf, type = 1, scale = FALSE)

```

Predicted CKD risk across alcohol quantity levels, were viewed separately by sex. A 2-way partial dependence plot of alcohol quantity and sex was mapped simultaneously against the variable CKD.

```{r}

pd <- partial(rf, pred.var = c("SEX_factor", "alcohol_quantity"), grid.resolution = 40, chull = TRUE)

autoplot(pd, contour = TRUE) +
  labs(title = "2D Partial Dependence Plot for Random Forest",
       x = "Sex",
       y = "Alcohol quantity in last week",
       fill = "Partial\nDependence") +
  theme_minimal()
```

The effect of alcohol quantity changes, depending on sex. This does indicate an interaction effect between sex and alcohol consumption on CKD. However this is not a formal hypothesis test. 

```{r}
#Plot with highly-aesthetic arguments
rpart.plot(tree, 
           type = 2,               # Split labels below the nodes
           extra = 104,            # Displays class, probabilities, and % of obs
           shadow.col = "gray",    # Subtle drop shadow under boxes
           nn = TRUE,              # Display node numbers
           faclen = 0)              

ts_class_err <- 1 - confusionMatrix(rf_data$CKD,
                                    predict(rf,
                                    newdata=rf_data,
                                    type="class"))$byClass["Recall"]

ts_class_err[1:2]

# MeanDecreaseAccuracy = permutation importance

options(digits=2)
rf$importance[,1:4]
varImpPlot(rf) #Shows what variables contributed most to the random forest

#Vote matrix
rf$votes

```

